---
title: "Is There a Gender Pay Gap?"
author: "Joy Xu"
date: "2026-10-09"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)

## Analysis
```
The analysis examined whether the population mean hourly gender pay gap differed from zero. Positive values indicate that women's mean hourly pay is lower than men's mean hourly pay.

```{r load-data}
library(tidyverse)
pay_gap <- read_csv("UK_gender_pay_gap_2017-18.csv")
```

## Distribution of the Gender Pay Gap

```{r figure, fig.width=8, fig.height=5}
ggplot(pay_gap, aes(x = DiffMeanHourlyPercent)) +
  geom_histogram(binwidth = 5) +
  coord_cartesian(xlim = c(-50, 100)) +
  labs(
    x = "Mean hourly gender pay gap (%)",
    y = "Number of employers",
    caption = "Figure 1. Distribution of the mean hourly gender pay gap. Display restricted to −50% to 100%; all observations were retained in the analysis."
  ) +
  theme_minimal()
```

## Results

A one-sample \(t\)-test indicated that the mean hourly gender pay gap was significantly greater than zero, \(M = 14.29\%\), 95% CI \([14.00\%, 14.58\%]\), \(t(10,531) = 96.50\), \(p < .0001\). On average, women's mean hourly pay was approximately 14.29% lower than men's mean hourly pay. Therefore, the null hypothesis that the population mean hourly gender pay gap equals zero was rejected.