---
title: "Is There a Gender Pay Gap?"
author: "Joy Xu"
date: "2026-10-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
## Load packages and data

```{r load-data}
library(tidyverse)

pay_gap <- read_csv("UK_gender_pay_gap_2017-18.csv")
```
## Inspect the data

```{r inspect-data}
glimpse(pay_gap)
```
## Data dictionary

Each row represents one employer's gender pay gap submission.

| Variable | Type | Description |
|:---|:---|:---|
| `EmployerName` | Character | Name of the employer submitting the gender pay gap report. |
| `DiffMeanHourlyPercent` | Numeric | Percentage difference between men's and women's mean hourly pay, calculated as \((\text{male mean} - \text{female mean}) / \text{male mean} \times 100\). A positive value indicates that women's mean hourly pay is lower; a negative value indicates that women's mean hourly pay is higher. |
## Mean gender pay gap

`DiffMeanHourlyPercent` measures the percentage difference between men's and women's mean hourly pay. Positive values indicate that men's mean hourly pay is higher.

```{r mean-pay-gap}
pay_gap |>
  summarise(
    mean_pay_gap = mean(DiffMeanHourlyPercent, na.rm = TRUE)
  )
```

## Distribution of the gender pay gap

```{r pay-gap-histogram}
ggplot(pay_gap, aes(x = DiffMeanHourlyPercent)) +
  geom_histogram(binwidth = 5) +
  coord_cartesian(xlim = c(-50, 100)) +
  labs(
    title = "Distribution of the Mean Hourly Gender Pay Gap",
    x = "Mean hourly gender pay gap (%)",
    y = "Number of employers",
    caption = "UK gender pay gap reporting data, 2017–18. Display restricted to −50% to 100%; all observations retained in the analysis."
  ) +
  theme_minimal()
```

## Hypothesis

The null hypothesis is that the population mean hourly gender pay gap is zero:

\(H_0: \mu = 0\)

The alternative hypothesis is that the population mean hourly gender pay gap is not zero:

\(H_1: \mu \neq 0\)

## One-sample t-test

```{r one-sample-t-test}
t.test(x = pay_gap$DiffMeanHourlyPercent, mu = 0)
```
## Results

A one-sample \(t\)-test indicated that the mean hourly gender pay gap was significantly greater than zero, \(M = 14.29\%\), 95% CI \([14.00\%, 14.58\%]\), \(t(10,531) = 96.50\), \(p < .0001\). On average, women's mean hourly pay was approximately 14.29% lower than men's mean hourly pay.